They are Very Smart: Genetic-Algorithm-Based Pathfinding in Games

### 2018-12-11 paper.tex

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned onslaught of enemies such as the fearsome Zombunny (see Figure 3). Zombunnies are driven by a single goal: destruction of the player's home base. Zombunnies spawn in groups of ten (separated by a slight delay) at the beginning of the game. Subsequent waves spawn after all members from the previous wave have died. Zombunnies use a genetic algorithm in conjunction with a recurrent neural network to evolve a pathfinding algorithm between generations (waves), with the long-term incentive of dealing damage to the player's base (explain fitness criteria more concisely here). Zombunnies can be killed by either taking damage from a player's defenses or through "starvation." A Zombunny will starve to death if it does not take or deal damage for a given length of time. Through this starvation mechanism, the Zombunnies which survive the longest are ultimately those which interact the most with the player's defenses, and are subsequently more likely to pass on attributes of their genome to the next generation. Zombunnies have three characteristic attributes: health, damage, and movement speed. All three of these stats are assigned upon spawning as dictated by their genome. Pathfinding is controlled by a recurrent neural net, whose weights are determined through a genetic algorithm between waves, which optimizes for Zombunnies' ability to achieve their overarching goal of player base destruction. Zombunnies can "see" nearby entities within a certain radius, including other Zombunnies, player defenses, and the player's home base (when in range). Based on the combination of visible inputs, Zombunnies will output a chosen direction which they will then walk in. This algorithm updates on a tick-by-tick basis. Zombunnies will damage structures they collide with. Because evolution of the pathfinding neural net only happens between generations, near the beginning of the game, Zombunnies will start off with relatively stupid pathfinding which will typically cause the first few populations to starve to death without dealing any significant damage. This provides the player with an initial grace period during which they may build their initial defenses. As time goes on, Zombunnies will evolve towards an optimal stat distribution of health, damage, and speed, as well as in their ability to pathfind in such a way as to thwart the player's defenses and end the game.

@@

Figure 18 shows the part of the FixedUpdate method which keeps track of Zombunny
death, destroying the Zombunny GameObject at the appropriate time and calling
the relevant death animations. For every tick that a Zombunny is not dead,
FixedUpdate will increment timeAlive, which is later used as a factor in
calculating fitness.

@@

The ZombieAttacker class calls Update on each tick, dealing damage to any item
in toAttack, which is a HashSet of Targetable GameObjects that intersect with
a Zombunny's attack collider (see Figure 4). This way, a Zombunny can attack
multiple GameObjects so long as they intersect with its attack collider. Damage
is applied to items in toAttack by calling their individual takeDamage methods,
with the damage dealt as determined by the Zombunny's genome passed as a
parameter.

\imfig{}{}

>>>>

\imfig{BaseExample}{An example set of player defenses around the home base}
In \tavs, the player must defend their home base against the aforementioned
onslaught of enemies such as the fearsome Zombunny (see Figure 3). Zombunnies
are driven by a single goal: destruction of the player's home base. Zombunnies
spawn in groups of ten (separated by a slight delay) at the beginning of the
game. Subsequent waves spawn after all members from the previous wave have died.
Zombunnies use a genetic algorithm in conjunction with a recurrent neural
network to evolve a pathfinding algorithm between generations (waves), with the
long-term incentive of dealing damage to the player's base% (explain fitness
% criteria more concisely here)
. Zombunnies can be killed by either taking damage
from a player's defenses or through "starvation." A Zombunny will starve to
death if it does not take or deal damage for a given length of time. Through
this starvation mechanism, the Zombunnies which survive the longest are
ultimately those which interact the most with the player's defenses, and are
subsequently more likely to pass on attributes of their genome to the next
generation. Zombunnies have three characteristic attributes: health, damage, and
movement speed. All three of these stats are assigned upon spawning as dictated
by their genome. Pathfinding is controlled by a recurrent neural net, whose
weights are determined through a genetic algorithm between waves, which
optimizes for Zombunnies' ability to achieve their overarching goal of player
base destruction. Zombunnies can "see" nearby entities within a certain radius,
including other Zombunnies, player defenses, and the player's home base (when in
range). Based on the combination of visible inputs, Zombunnies will output a
chosen direction which they will then walk in. This algorithm updates on a
tick-by-tick basis. Zombunnies will damage structures they collide with. Because
evolution of the pathfinding neural net only happens between generations, near
the beginning of the game, Zombunnies will start off with relatively stupid
pathfinding which will typically cause the first few populations to starve to
death without dealing any significant damage. This provides the player with an
initial grace period during which they may build their initial defenses. As time
goes on, Zombunnies will evolve towards an optimal stat distribution of health,
damage, and speed, as well as in their ability to pathfind in such a way as to
thwart the player's defenses and end the game.

@@

Figure 18 shows the part of the FixedUpdate method which keeps track of Zombunny
death, destroying the Zombunny GameObject at the appropriate time and calling
the relevant death animations. For every tick that a Zombunny is not dead,
FixedUpdate increments timeAlive, which is later used as a factor in
calculating fitness.

@@

The ZombieAttacker class (see Figure 19) calls Update on each tick, dealing damage to any item
in toAttack, which is a HashSet of Targetable GameObjects that intersect with
a Zombunny's attack collider (see Figure 4). This way, a Zombunny can attack
multiple GameObjects so long as they intersect with its attack collider. Damage
is applied to items in toAttack by calling their individual takeDamage methods,
with the damage dealt as determined by the Zombunny's genome passed as a
parameter.

\imfig{ZombieAttacker}{The Update method of the ZombieAttacker class}
